---
title:  "Convergence Analysis"
author: "Helen Ristov, Jing Song, and Ryan Ranson"
date:   "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
---

```{r preparation, echo=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
library(plyr)
library(xtable)

load("PreprocessedConvergence.RData")

res.summary <- ddply(res, .(perf.period, threshold), function(x) data.frame( 
                     obs           = sum(x$obs), 
                     disjoint.obs  = sum(x$disjoint.obs),
                     converged.obs = sum(x$converged.obs),
                     disjoint.pct  = sum(x$disjoint.obs) / sum(x$obs),
                     converged.pct = sum(x$converged.obs) / sum(x$disjoint.obs),
                     avg.reversion = weighted.mean(x$avg.reversion, x$disjoint.obs, na.rm = TRUE),
                     avg.shortfall = weighted.mean(x$avg.shortfall, x$disjoint.obs, na.rm = TRUE)))

df.metrics <- ddply(res, ~Series, summarise, 
                    obs  = mean(obs), 
                    mean = mean(mean),
                    sd   = mean(std.dev), 
                    min  = mean(min),
                    max  = mean(max))

df.summary <- data.frame(Mean    = mean(df.metrics$mean),
                         Std.Dev = mean(df.metrics$sd),
                         Min     = mean(df.metrics$min),
                         Max     = mean(df.metrics$max))
```

## Introduction

The goal of this analysis is to identify points in double fly price series where there is
a high probability of mean reversion and also to determine an appropriate expected holding 
period for trades.

Definitions

- Limit: assumed long-term mean of the double fly price series
- Epsilon: distance above and below the limit which defines the convergence zone
- Threshold: distance above and below the limit for which we will consider an observation disjoint
- Performance Period: time period we will look at each disjoint instance to test for convergence

A disjoint observation is said to converge if it reaches the convergence zone at any point
during the performance period. The reversion metric is the change in the direction of the limit
of the series until either the series converges or reaches the end of the performance period.

We will look at 60 CL double fly prices from CLF0.G0.H0.J0 to CLZ4.F5.G5.H5. The limit
parameter is assumed to be 0 and epsilon is set at 0.01.

## Data Summary

```{r, echo=FALSE}
ggplot(data = res, aes(x = Series, y = mean, ymin = min, ymax = max)) + 
  geom_crossbar(fill = "lightblue", color = "black") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  ylab("Range")
```

## Convergence Metrics {.columns-2}

- Convergence tends to increase with performance period
- The relationship between convergence rates and threshold is less clear
- High convergence rates for high threshold and performance periods is in part to due small sample size (there are only `r res.summary$disjoint.obs[res.summary$perf.period == 30 & res.summary$threshold == 0.25]` disjoint observations for performance period = 30 and threshold = 0.25)
  
```{r, echo=FALSE, fig.width=14, fig.height=10}
ggplot(res.summary, aes(x = perf.period, y = threshold)) +
  geom_tile(aes(fill = avg.reversion), color = "white") +
  scale_fill_gradient(low = "red", high = "green") + 
  xlab("Performance Period") +
  ylab("Threshold") + 
  labs(fill = "Reversion")
```

## Reversion Metrics

```{r, echo=FALSE}
plot(cars)
```

